Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-07-31 articles/formalism/sections/introduction.tex
Minor cleanups in introduction.

The dominant contribution to the $E$-modes of polarization is sourced by primordial scalar perturbations, but these do not generate B-modes of polarization at first order. Various phenomena may generate $B$-modes in polarization measurements of the microwave sky:
% Need citations for all...
primordial tensor perturbations (gravitational waves)%, which reveal the energy scale of inflation
\citep{1997PhRvD..56..596H,1997PhRvL..78.2054S};
weak gravitational lensing of $E$-modes induced by the potentials of intervening large scale structures;
Milky Way foregrounds (especially Galactic synchrotron and dust emission)
\citep{2016A&A...586A.133P};
uncorrected systematic problems in the data \citep{2003PhRvD..67d3004H,2008PhRvD..77h3003S};
and unknown, exotic phenemena like cosmic birefringence or primordial magnetic fields
\citep{1996ApJ...469....1K,1999PhRvL..83.1506L,2004ApJ...616....1C,2014MNRAS.438.2508P}.

@@

The primary aim of the current CMB experiments is to make precise measurements of the CMB polarizations. While the $E$-modes of polarization have been measured reasonably well by a number of experiments (\cite{2018RPPh...81d4901S} gives a recent review), the ongoing and future CMB experiments aim to measure the B-modes of CMB polarization with unprecedented accuracy. The instruments are swiftly approaching the desired sensitivities to enable us to in principle measure $B$-mode signals with $r\gtrsim 0.001$ \cite{Spider, CLASS,Litebird, BICEP22015, 2016arXiv161002743A, 2017arXiv170602464A,Remazeilles2018}. However $B$-modes generated by galactic foreground are expected to be a few orders of magnitude higher than $B$-mode amplitude measurable by these detectors.  Precise modelling and subtraction of this large foreground contribution poses a major challenge for a robust unravelling of the minuscule $B$-mode signal.

The formalism for converting the Stokes parameters to scalar quantities is well established \citep{1997PhRvD..55.7368K,1997PhRvD..55.1830Z}. The spin-0 scalar $E$/$B$ modes relate to the spin-2 complex Stokes parameters via the spin-raising and -lowering operators ($\eth^2,\bar \eth^2$), which are derivatives evaluated locally and filtered.  In practice the $E$/$B$ modes are computed to some specified band limit, and the filtering and band limit make them non-local functions of the polarization field.  In other words, $E$/$B$ modes evaluated at a point receive contributions from all over the sky. In this work we aim to gain real space insights into the non-locality of the $E/B$ fields compared to the Stokes parameters. With renewed focus on foreground contamination to the $B$-mode signal we aim to gain an intuition for $E/B$ mode patterns resulting from physical polarized structures in the galaxy. These real space insights may yield new ideas for minimizing foreground contamination which may not be obvious using conventional approaches.

>>>>

The dominant contribution to the $E$-modes of polarization is sourced by primordial scalar perturbations, but these do not generate $B$-modes of polarization at first order. Various phenomena may generate $B$-modes in polarization measurements of the microwave sky:
% Need citations for all...
primordial tensor perturbations (gravitational waves)%, which reveal the energy scale of inflation
\citep{1997PhRvD..56..596H,1997PhRvL..78.2054S};
weak gravitational lensing of $E$-modes induced by the potentials of intervening large scale structures;
Milky Way foregrounds (especially Galactic synchrotron and dust emission)
\citep{2016A&A...586A.133P};
uncorrected systematic problems in the data \citep{2003PhRvD..67d3004H,2008PhRvD..77h3003S};
and unknown, exotic phenemena like cosmic birefringence or primordial magnetic fields
\citep{1996ApJ...469....1K,1999PhRvL..83.1506L,2004ApJ...616....1C,2014MNRAS.438.2508P}.

@@

The primary aim of the current CMB experiments is to make precise measurements of CMB polarization. While the $E$-modes of polarization have been measured reasonably well by a number of experiments (\cite{2018RPPh...81d4901S} gives a recent review),  ongoing and future CMB experiments aim to measure the $B$-modes of CMB polarization with unprecedented accuracy. The instruments are swiftly approaching the desired sensitivities to enable us to in principle measure $B$-mode signals with $r\gtrsim 0.001$ \cite{Spider, CLASS,Litebird, BICEP22015, 2016arXiv161002743A, 2017arXiv170602464A,Remazeilles2018}. However $B$-modes generated by galactic foreground are expected to be a few orders of magnitude higher than $B$-mode amplitude measurable by these detectors.  Precise modelling and subtraction of this large foreground contribution poses a major challenge for a robust unravelling of the minuscule $B$-mode signal.

The formalism for converting the Stokes parameters to scalar quantities is well established \citep{1997PhRvD..55.7368K,1997PhRvD..55.1830Z}. The spin-0 scalar $E$/$B$ modes relate to the spin-2 complex Stokes parameters via the spin-raising and -lowering operators ($\eth^2,\bar \eth^2$), which are derivatives evaluated locally and filtered.  In practice the $E$/$B$ modes are computed to some specified band limit, and the filtering and band limit make them non-local functions of the polarization field.  In other words, $E$/$B$ modes evaluated at a point receive contributions from all over the sky. In this work we aim to gain real space insights into the non-locality of the $E/B$ fields compared to the Stokes parameters. With renewed focus on foreground contamination to the $B$-mode signal we aim to gain an intuition for $E/B$ mode patterns resulting from physical polarized structures in the galaxy. These real space insights may yield new ideas for minimizing foreground contamination that are not obvious using conventional approaches.
